The paper on blockchain front-running

### 2018-09-25 frontrunning/Latex/sections/paper.tex
Moving to ICO (everything in blue needs more explanation)

 that is well-connected front-run other transactions in the network. For regular users to front-run others on the blockchain, they need to be well connected to other nodes on the network. Doing so, they are able to listen to the network and monitor all transactions that are broadcast. On the Ethereum blockchain, users have to pay for the computations in a small amount of Ether called \textit{gas}~\cite{AccountT67:online}. The price that users pay for transactions, \textit{gasPrice}, can increase or decrease how quickly miners will execute them and include them within the blocks they mine. Once seeing two identical transactions with different transaction fees, a rational miner will prioritize the transaction that pays a higher gas price per unit of gas --gasPrice--, this is due to limited space in the blocks while maximizes their profit in the block they mine. Therefore, any regular users who run a full-node Ethereum client can front-run pending transactions by sending similar transaction with higher gas price. Blockchain miners are the only parties who can decide on the order of transactions within a block they mine, they can easily intercept and reorder the transactions in their blocks, this in case of malicious reordering is known as \textbf{transaction reordering} attack.

@@

Aune~\etal discuss how the lack of time priority between broadcasting a transaction and its validation by miners on a blockchain based system would lead to the market information leakage~\cite{aune2017footprints}.They also propose a cryptographic approach, similar to commit and reveal (see Section~\ref{CommitmentScheme}) to prevent front-running. Malinova and Park discuss different design settings for financial markets and trading platforms on decentralized ledgers~\cite{malinova2017market}. They argue that blockchain brings new options of transparency which affects traders behaviour, especially in the presence of intermediary front-runners. They model indirect trading costs and use the economic literature on search and trading at decentralized exchanges in their design settings to make front-running costly and inefficient. Breidenbach~\etal~\cite{breidenbach2018enter} face the front-running issue in the context of automating bug bounties for smart contracts. Upon submitting the bug bounty to the Hydra smart contract, one can front-run the bug report and claim the bounty instead of the original reporter. In order to mitigate this issue they proposed \textit{Submarine Commitments} which extend commit and reveal; we discuss them further in Section~\ref{submarineCommits}.

@@

\subsection{Markets and Exchanges.} \label{sec:frontrunningExchanges}

@@

%Tragedy of the commons
The first category of DApp in Table~\ref{tab:top25DApps} is exchanges. 
Exchanges such as EtherDelta\footnote{also known as ForkDelta for the UI \url{https://forkdelta.app/}}, is the closest implementation to decentral exchanges, however their order books are stored on a central server they control and shown on their user interface (website). Other than the fact that they can re-order or censor the orders on their servers, there are other risks of front-running nature in these designs. Similarly 0xProtocol~\cite{warren20170x} uses \textit{Relayers} which act as the order book holders and are not prone to cheating. One main issue with these designs is the \textit{cancellation grief}. Although the order books are not in the smart contract, in order to prevent denial of service attacks, the user needs to send a transaction to cancel her order. In this case, when an adversarial actor sees the cancellation transaction, he sends a fill order transactions with higher gasPrice to get in front of the cancellation order and take the order before it is canceled. As mentioned in ~\ref{minersfrontrunning}, miners are in a better position to perform such an attack.

@@

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.7\linewidth]{figures/Miner_reorder_frontrun.png}
\caption{The adversarial miner can monitor the Ethereum mempool for decentral exchange cancel orders and upon seeing the cancellation transaction, he puts his buy order prior to the cancel transaction. Doing so, the miner can profit from the underlying trade and also get the gas included in the cancel transaction. \label{fig:MinerFrontrunning}} 
\end{figure}
%https://www.lucidchart.com/invitations/accept/c8768786-4a3f-4c78-b3c8-aa04a6b81d06

see figure~\ref{fig:RegularFrontrunning}. For example, in a decentral exchange, he can monitor the network and once seeing a pending \textit{buy} transaction which will further increase the price of the asset, a front-runner can send a transaction with a higher gas price and hope to have his transaction mined ahead of the original pending transaction. By doing so, he achieves a better rate from the buyer. Note that in this case, blockchain front-runners are visible to the network as they broadcast their transactions to all the network participants. The probability of success in such attach can be increased by flooding the network by transactions with higher than average gasPrice.

For example in an Ethereum based application, a miner learns about the pending \textit{buy} transaction of 1000 units of a token\ie TKN, presumably if this transaction goes through, it causes the price of the purchased token to increase. So the dishonest miner can step in front of this transaction and place his own buy order ahead. He would simply create his \textit{buy} of 1000 TKN and include it in front of the previous \textit{buy} transaction in the block he mines. Doing so, he would receive a better rate than other network users, and afterward can sell the assets he has bought and gained a price advantage. It should be noted that miners can only front-run transactions within the blocks they happen to mine. They could do this also by not broadcasting their own transactions to the network before including it in the block, this makes the miners be less noticeable to the network when front-running.

Another example here is in the case of decentral exchanges and order cancellation. Every exchange should have the functionality to cancel the orders, especially for a volatile market. In this case, when the user decides that her order is not profitable anymore, she would send a cancelation transaction. Here the miner sees the cancelation transaction and puts his own transaction to fill that order in front of the cancellation transaction --transaction reordering--, potentially profiting from the order. He can also include the canceled transaction in the block to collect the full gas limit used by the cancel transaction~\cite{CostofDecentralization:online}, see figure~\ref{fig:MinerFrontrunning}. Front-running can also occur in non-financial smart contracts. As an example, a smart contract which adds all the participants to a party invite list could only be closed by the smart contract owner. In this case, when the owner sends the transaction to close the list, a miner can include his own list of participants before the close transaction in the block he mines. This is an issue with the design of the smart contracts and is known as transaction-ordering dependence vulnerability~\cite{luu2016making}.

Bancor is another Ethereum-based application that allows users to exchange their tokens without any counter-party risk. This protocol aims to solve the cryptocurrency liquidity issue by introducing \textit{Smart Tokens}~\cite{hertzog2017bancor}-- ERC20 compatible tokens with a built-in liquidity mechanism that is always available to users. ERC20 is the interface standard for implementing tokens on top of Ethereum~\cite{erc20}. Smart Tokens can be bought and sold through the users smart contract at an automatically calculated price which displays supply and demand. Doing so, Bancor protocol provides continuous liquidity for digital assets without relying on an order book as there is no requirement to match sellers and buyers.
Implemented on the Ethereum blockchain, when transactions are broadcast to the network, they sit in a pending transaction pool known as \textit{mempool} waiting for the miners to mine them. Since Bancor handles all the trades and exchanges on the Ethereum blockchain (unlike other existing decentralized exchanges), these transactions are all visible to the public for some time before being included within a block. This leaves this blockchain-based decentralized exchange vulnerable to the blockchain race condition attack as attackers are given enough time to front-run other transactions and gain favorable profits~\cite{BancorIs7:online}. Researchers have shown and implemented a proof of concept code to front-run Bancor as a regular non-miner user~\cite{NewTab13:online}.

%______________________________________________________________%
\subsection{Crypto-Collectibles Games.}

The second category of DApp in Table~\ref{tab:top25DApps} is crypto-collectables.

Crypto-Collectibles, also known as \textit{ERC-721 Non-Fungible Token Standard}~\cite{erc721}, is one of the applications of Ethereum blockchain. Cryptokitties are known to be one of the first and the most popular use cases of ERC721 tokens~\cite{cryptokitties}. Their market cap reached more than 6 million dollars in the first few months their launch, however, declined soon after. Crypto-Collectibles, are similar to ERC-20 tokens and can be listed on exchanges. Most of the front-running issues applied to the exchanges also apply to DApps using ERC-721 tokens. \\

As an example Cryptokitties can be bred, bought or sold through the main interface\footnote{Cryptokitties website \url{https://www.cryptokitties.co/}}. %see Figure~\ref{fig:cryptokittie}.
To buy a cryptokittie, the user sends a bid transaction \textit{bid(uint256 \_tokenId)} containing the cryptokittties ID which can be found either on their website or on the ERC721 token smart contract. This is similar to open auctions, bid value and the object bidding on is visible to the network and any user could be easily front-run by sending the same transaction with higher gasPrice to replace the initial bidder. 
% https://etherscan.io/tx/0x62a76b7e345e8abd463487d0348c118a9314b197024e090fee7b1fbce434da40

%Possibly remove this figure due to page limit
\begin{figure}[h]
\centering
\includegraphics[width=0.3\linewidth]{figures/cryptokittie842912.png}
\caption{ Cryptokitty Number 842912 \label{fig:cryptokittie}}
\end{figure}

%______________________________________________________________%

@@

The third category of DApp in Table~\ref{tab:top25DApps} is gambling services.

Due to the deterministic nature of blockchains, running gambling games based on randomness is close to impossible. There had been attempts to use block headers or such as the seed of randomness but almost all have failed as every data accessible to the smart contract at the time of execution, is also accessible to other smart contracts and they could predict the output on the run, thus prone to front-running and miner\'s manipulation. Many infamous gambling games are now running on Ethereum network and front-running them requires a level of sophistication and knowledge of the internal of the games and Ethereum.\\
One of the most popular smart contracts on Ethereum blockchain belongs to a controversial gambling game named \texttt{Fomo3d}, also known as \textit{Exit Scam}\footnote{\url{https://exitscam.me/play}}. The aim of this game is to be the last person purchasing the key when the timer goes to zero in which every ticket purchase increases the timer by 30 seconds. It was believed that because of the nature of the game, it will never end, however on August 22, 2018 the first round of the game ended and the winner taking 10,469 Ether\footnote{First winner of Fomo3D, won 10,469 Ether \url{ https://etherscan.io/tx/0xe08a519c03cb0aed0e04b33104112d65fa1d3a48cd3aeab65f047b2abce9d508}}, equivalent to 2.1 Million dollars on the time of writing. 
It is believed that the winner, used a sophisticated approach to finish last~\cite{fomo3dhacker}, we call this approach \textit{Denial-of-service Front-running}. The winner deployed a few high gas consumption smart contracts prior to the win when the timer of the game reached about 3 minutes, the winner bought 1 ticket and sent multiple high gasPrice transactions to aforementioned smart contracts to lure the miners to mine these highly profitable transactions. For a DoS front-running to work, the attacker floods the network with high gasPrice transactions to prevent other transactions to go through. Luckily enough the winner was able to congest the network in a way that no other ticket purchases were mined in the 3 minutes window and the miners were prioritizing the high gasPriced transactions over other transactions. All the details mentioned here are drawn from the transaction history on the blockchain and unless otherwise proven seem to be the way the winner acted to win the pot. 
%probably needs more work to be good enough. more resources will be out in the coming days!

%______________________________________________________________%
\subsection{Name Services.}\label{sec:ens}

The final category of DApp in Table~\ref{tab:top25DApps} is name services.

Front-running can also occur within non-financial blockchain applications such as naming systems. Recently, many blockchain-based domain registrars have been designed and implemented to eliminate the role of central parties \ie domain name system (DNS) which itself introduced the single point of failures to the entire web. Ethereum Name Service (ENS)\footnote{\url{https://ens.domains/}}, is the most popular decentral naming service that uses a sealed auction to facilitate the procedures of purchasing new \texttt{.eth} domain names. Although sealed auction method is used in ENS implementation, the normal workflow uses \textit{startAuctionsAndBid()} with the deposit equal or higher than the bid, this information leakage could be used to front-run the transaction with the same bid amount or slightly higher to outbid the initial sealed bid. By doing so the front-runner invalidates the original bid and hence registers the domain name himself. Ghazal is another Ethereum-based naming system that allows users to register \texttt{.ghazal} domain names and bind certificates to those names~\cite{moosavighazal}.  As Ghazal is using a registration method in plain text, it is susceptible to front-running. 
Note that front-running domain names has been a subject of discussion outside the context of blockchian applications before~\cite{sac022en33:online}. Domain name registrars can steal available domain names upon seeing a query from an interested user for the availability of the domain. In this case, the registrar can register the domain name ahead and sell it back to the user for a higher price.

%% = = = = = = = = = = = = = = = = = = = = = = = = = = %
% = = = ICOs
% = = = = = = = = = = = = = = = = = = = = = = = = = = %%

@@

>>>>

 that is well-connected front-run other transactions in the network. For regular users to front-run others on the blockchain, they need to be well connected to other nodes on the network. Doing so, they are able to listen to the network and monitor all transactions that are broadcast. On the Ethereum blockchain, users have to pay for the computations in a small amount of Ether called \textit{gas}~\cite{AccountT67:online}. The price that users pay for transactions, \textit{gasPrice}, can increase or decrease how quickly miners will execute them and include them within the blocks they mine. Once seeing two identical transactions with different transaction fees, a rational miner will prioritize the transaction that pays a higher gas price per unit of gas, \textsf{gasPrice}, due to limited space in the blocks. Therefore, any regular users who run a full-node Ethereum client can front-run pending transactions by sending similar transaction with higher gas price. Blockchain miners are the only parties who can decide on the order of transactions within a block they mine, they can easily intercept and reorder the transactions in their blocks, this in case of malicious reordering is known as \textbf{transaction reordering} attack.

@@

Aune~\etal discuss how the lack of time priority between broadcasting a transaction and its validation by miners on a blockchain based system would lead to market information leakage~\cite{aune2017footprints}.They also propose a cryptographic approach, similar to commit and reveal (see Section~\ref{CommitmentScheme}) to prevent front-running. Malinova and Park discuss different design settings for financial markets and trading platforms on decentralized ledgers~\cite{malinova2017market}. They argue that blockchain brings new options of transparency which affects traders behaviour, especially in the presence of intermediary front-runners. They model indirect trading costs and use the economic literature on search and trading at decentralized exchanges in their design settings to make front-running costly and inefficient. Breidenbach~\etal~\cite{breidenbach2018enter} face the front-running issue in the context of automating bug bounties for smart contracts. Upon submitting the bug bounty to the Hydra smart contract, one can front-run the bug report and claim the bounty instead of the original reporter. In order to mitigate this issue they proposed \textit{Submarine Commitments} which extend commit and reveal; we discuss them further in Section~\ref{submarineCommits}.

@@

% = = =

\subsection{Markets and Exchanges} \label{sec:frontrunningExchanges}

@@

%Tragedy of the commons

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.7\linewidth]{figures/Miner_reorder_frontrun.png}
\caption{The adversarial miner can monitor the Ethereum mempool for decentralized exchange cancel orders and upon seeing the cancellation transaction, he puts his buy order prior to the cancel transaction. Doing so, the miner can profit from the underlying trade and also get the gas included in the cancel transaction. \label{fig:MinerFrontrunning}} 
\end{figure}
%https://www.lucidchart.com/invitations/accept/c8768786-4a3f-4c78-b3c8-aa04a6b81d06

The first category of DApp in Table~\ref{tab:top25DApps} is exchanges. Exchanges such as EtherDelta\footnote{also known as ForkDelta for the UI \url{https://forkdelta.app/}}, purport to implement a decentralized exchange, however their order books are stored on a central server they control and shown on their users with a website interface. Central exchanges can front-run orders in the traditional sense, as well as re-order or block orders on their servers. 0xProtocol~\cite{warren20170x} uses \textit{Relayers} which act as the order book holders \textblue{and are not prone to cheating.} One main issue with the relay design is \textit{cancellation grief}. To \textblue{prevent denial of service attacks}, the user needs to send a transaction to cancel her order. In this case, an adversarial actor many see a pending  cancellation transaction, and send  a fill order transaction with higher gasPrice to get in front of the cancellation order and take the order before it is canceled. This is illustrated in Figure~\ref{fig:MinerFrontrunning}.

@@

Designing truly decentralized exchanges, where the orderbook is implemented directly on a public blockchain, is being pursued by a number of projects [\textblue{cite github page}]. These designs generally face the following attack (illustrated in Figure~\ref{fig:RegularFrontrunning}). An adversary can monitor the network with a full node for pending \textit{buy} (or \textit{sell}) transactions which could increase (or \textit{decrease}) the future price of the asset. A front-runner can send a competing orders with a higher gas price, and hope to have her transactions mined ahead of the original pending transaction. The benefits of doing this vary. If the original bid is a large market order (\ie it will execute at any price), the adversary can front-run a pair of limit orders that will bid near the best offer price and offer at a higher price. If these execute ahead of the market order, the front-runner profits by scalping the price of the shares. If the adversary has pre-existing offers likely to be reached by the market order, she could front-run a cancelation and a new offer at a higher price. A simpler attack would be to guage the demand for trades at a given price by the number of pending orders, and to front-run at the same price in hopes that the market demand is the result of accurate new market information. Miners are in the best position to conduct these attacks as they hold fine-grained control over the exact set of transactions that will execute and in what order and can mix in their own (late) transactions without broadcasting them. Miners do however have to commit to what their own transactions will be before beginning the proof of work required to solve a block.

% Front-running can also occur in non-financial smart contracts. As an example, a smart contract which adds all the participants to a party invite list could only be closed by the smart contract owner. In this case, when the owner sends the transaction to close the list, a miner can include his own list of participants before the close transaction in the block he mines. This is an issue with the design of the smart contracts and is known as transaction-ordering dependence vulnerability~\cite{luu2016making}.

Bancor is another Ethereum-based application that allows users to exchange their tokens without any counter-party risk. This protocol aims to solve the cryptocurrency liquidity issue by introducing \textit{Smart Tokens}~\cite{hertzog2017bancor}. Smart tokens are ERC20-compatible and can be bought or sold through a DApp-based dealer that is always available and implements a market scoring rule to manage its prices. Doing so, Bancor provides continuous liquidity for digital assets without relying on brokers to match buyers with sellers. Implemented on the Ethereum blockchain, when transactions are broadcast to the network, they sit in a pending transaction pool known as \textit{mempool} waiting for the miners to mine them. Since Bancor handles all the trades and exchanges on the Ethereum blockchain (unlike other existing decentralized exchanges), these transactions are all visible to the public for some time before being included within a block. This leaves this blockchain-based decentralized exchange vulnerable to the blockchain race condition attack as attackers are given enough time to front-run other transactions \textblue{and gain favorable profits}~\cite{BancorIs7:online}. Researchers have shown and implemented a proof of concept code to front-run Bancor as a regular non-miner user~\cite{NewTab13:online}.

% = = =

\subsection{Crypto-Collectibles Games}

The second category of DApp in Table~\ref{tab:top25DApps} is crypto-collectables. Consider Cryptokitties,\footnote{Cryptokitties website \url{https://www.cryptokitties.co/}} the most active DApp in this category and third most active overall. Each cryptokitty (see Appendix~\ref{app:cryptokitty}) is a cartoon kitten with a set of unique features to distinguish it from other cryptokitties, some features more rare and harder to obtain. They can be bought, sold, or bred with other cryptokitties. At the Ethereum level, the kitty is a token implemented with \textit{ERC-721: Non-Fungible Token Standard}~\cite{erc721}. The market cap of cryptokitty tokens peaked at more than 6 million dollars in the first few months their launch, however, has declined since. ERC-721 are similar enough to ERC-20 tokens that they can be listed on exchanges setup for ERC-20. To buy a cryptokittie, the user sends the following bid transaction: \texttt{bid(uint256 \_tokenId)}. It contains the cryptokittty's ID which can be found either on their website or on the ERC721 token smart contract. \textblue{This is similar to open auctions, bid value and the object bidding on is visible to the network and any user could be easily front-run by sending the same transaction with higher gasPrice to replace the initial bidder.}
% https://etherscan.io/tx/0x62a76b7e345e8abd463487d0348c118a9314b197024e090fee7b1fbce434da40

% The blue above needs expansion with further details.

% = = =

@@

The third category of DApp in Table~\ref{tab:top25DApps} is gambling services. While a large category of gambling games are based on random outcomes, DApps do not have unique access to an unpredictable data stream to harvest for randomness. Any candidate (such as block headers) source of randomness is accessible to all DApp functions and can also be manipulated to an extent by miners.

\textsf{Fomo3D} is an example of a game style (known as \textsf{Exit Scam}\footnote{\url{https://exitscam.me/play}}) not based on random outcomes, and it is the most active game on Ethereum in our sample. The aim of this game is to be the last person to have purchased a ticket when a timer goes to zero in a scenario where anyone can buy a ticket and each purchase increases the timer by 30 seconds. Many speculated such a game would never end but on August 22, 2018, the first round of the game ended with the winner collecting 10,469 Ether\footnote{The first winner of Fomo3D, won 10,469 Ether \url{ https://etherscan.io/tx/0xe08a519c03cb0aed0e04b33104112d65fa1d3a48cd3aeab65f047b2abce9d508}} equivalent to \$2.1M USD at the time. 
Blockchain forensics indicate a sophisticated winning strategy to frontrun new ticket purchases~\cite{fomo3dhacker} that would reset the counter. The winner appears to have started by deploying some high gas consumption DApps unrelated to the game. When the timer of the game reached about 3 minutes, the winner bought 1 ticket and then sent multiple high gasPrice transactions to her own DApps. These transactions congested the network and bribed miners to prioritize them ahead of any new ticket purchases in \textsf{Fomo3D}.

%probably needs more work to be good enough. more resources will be out in the coming days!

% Is this the same as order flooding in HFT?

% = = =

\subsection{Name Services.}\label{sec:ens}

The final category in Table~\ref{tab:top25DApps} is name services, which are primarily aimed at displacing central parties involved in web domain registration (\eg ICAAN and registrars) and resolution (\eg DNS). For simple name services (such as some academic work like Ghazal~\cite{moosavighazal}), domains purchases are transactions and network monitors can front-run other users attempting to register domains. This parallels front-running attacks seen in regular (non-blockchain) domain registration~\cite{sac022en33:online}. \textsf{Ethereum Name Service (ENS)}\footnote{\url{https://ens.domains/}} is the most active naming service on Ethereum. Instead of allowing new  \texttt{.eth} domain names to be purchased directly, they are put up for a sealed bid auction. \textblue{Although sealed auction method is used in ENS implementation, the normal workflow uses \texttt{startAuctionsAndBid()} with the deposit equal or higher than the bid, this information leakage could be used to front-run the transaction with the same bid amount or slightly higher to outbid the initial sealed bid. By doing so the front-runner invalidates the original bid and hence registers the domain name himself.}

%% = = = = = = = = = = = = = = = = = = = = = = = = = = %
% = = = ICOs
% = = = = = = = = = = = = = = = = = = = = = = = = = = = %

@@

\textblue{Cite Princeton paper on Namecoin for frontrunning mitigation using commit/reveal.}
